Metal Mixing and Ejection in Dwarf Galaxies is Dependent on Nucleosynthetic Source

### 2018-07-30 main.tex
updateing more references in introduction. significant clarification points from KVJ comments, particularly in interpretation and intro. Still have a couple more KVJ comments to address

One-zone chemical evolution models assume homoegeneous mixing of metals from recent star formation into gas available for future star formation \citep[e.g.][]{Lanfranchi2006b,Kirby2011-III,Cote2017,Andrews2017}. While this assumes that metal mixing in the ISM plays no role in delaying future enrichment in star formation, accounting for this process is challening. More complex models \citep[e.g.][]{SchonrichBinney2009,Pezzulli2016} employ multiple (usually radial) zones or follow multiple gas phases to attempt to account for these effects. How to properly account for multi-phase mixing, however, is still poorly understood. This is in large part because both these models and large-scale cosmological simulations lack the necessary fidelity to capture the detailed, multi-phase mixing process of metals in the ISM directly. Recent hydrodynamics simulations have employed parameteric models to account for unresolved sub-grid metal mixing \citep{PanScannapiecoScalo2013,Sarmento2017}, which plays an important role in determining the chemical properties of galaxies \citep[e.g.][]{Shen2010, Pilkington2012, Few2012, Brook2014, Escala2018}. Consequently, following metal mixing, if treated in these models, requires assumptions about how metals mix within the ISM and how metals couple to galactic outflows. When metals are tracked individually, as opposed to a global metallicity field, their injection, mixing, and outflow properties are often treated uniformly. However, these processes are critical components in understanding galactic chemical evolution. Detailed hydrodynamics simulations incorporating a realistic multi-phase ISM and detailed stellar feedback are required to understand metal mixing and metal outflows \citep[e.g.][]{FengKrumholz2014,Armillotta2018}. In addition, it remains to be seen to what degree, if at all, metals of different nucleosynthetic origins may couple differently to the ISM.

The use of low mass dwarf galaxies, both observationally in the Local Group and in theoretical models, has been critical in improving our understanding of galactic chemical evolution. That dwarf galaxies efficiently pollute the circumgalactic medium and intergalactic medium (IGM) with metals has been demonstrated for some time both theoretically \citep[e.g.][]{DekelSilk1986,MacLowFerrara1999,Fragile2004}, and with direct observational evidence from the metal retention fractions of Local Group dwarf galaxies \citep[e.g.][]{Kirby2011-metals,McQuinn2015}. It is commonly assumed that all metals couple equally to galactic outflows, however, which may not be a valid assumption if metals of different nucleosynthetic origins couple differently to the ISM \citep{KrumholzTing2018}. Relaxing this assumption has implications for both interpreting observations of stellar abundances in nearby dwarf galaxies and in modeling galactic chemical evolution in both semi-analytic models and lower resolution cosmological simulations. However, testing this assumption has only become possible recently as it requires high resolution, galaxy-scale hydrodynamics simulations that can resolve ISM mixing and self-consistently drive galactic winds through multi-channel stellar feedback.

It is becoming increasingly important to develop a concrete theoretical understanding of galactic chemical evolution as a result of multiple, recent observational campaigns to probe detailed stellar abundances in our Milky Way and the Local Group, such as SEGUE \citep{Yanny2009}, RAVE \citep{Kunder2017}, APOGEE \citep{Anders2014}, and GALAH \citep{Buder2018}. Stellar abundances are directly imprinted with the enrichment pattern of their star forming cloud, whose chemical properties are determined by the process of turbulent metal mixing in the ISM. The degree to which we can associate "chemically tagged" stars as co-eval depends directly upon our understanding of metal enrichment and metal mixing in the ISM. However, this understanding is critical for using these observations to further deduce properties of a galaxy's evolutionary history.

@@

\subsection{Physical Interpretation of the PDF}
\label{sec:interpretation}
Take the simple case of an initially primordial, uniform, isothermal medium of mass $M_o$ with initial metallicity $Z_o = 0$, containing a single, un-mixed enrichment event of mass $M_{\rm ej}$ whose size is small compared to the system and mass $M_{\rm ej} / M_o \ll 1$. Thus the background medium represents a virtually inexhaustible (but finite) source of un-enriched gas. In this case, $p(Z)$ takes the form of a double-delta function
\begin{equation}
p(Z) = \delta(Z) + \frac{M_{\rm ej}}{M_o}\delta(Z - Z_{\rm ej}).
\end{equation}
If the enriched material mixes continually with the primordial gas at a constant rate, after some time this gas will have mixed with an equal amount of primordial gas. At this point, the high metal fraction delta term becomes $\frac{2M_{\rm ej}}{M_o} \delta(Z - \frac{1}{2}Z_{\rm ej})$. This implies an inverse relationship between the mass of enriched gas and the metallicity of the enriched gas, and thus a power-law evolution in time of the high metallicity portion of $p(Z)$ with slope $\alpha = 1$. If the gas is continually enriched by identical, well-separated enrichment events of mass $M_{\rm ej}$ and metallicity $Z_{\rm ej}$, $p(Z)$ will become a power-law in $Z$, truncated at some time dependent minimum $Z$. The slope of the power-law is determined by the rate of injection versus mixing with the ambient medium. A power-law index $\alpha < 1$ can occur when injection occurs more rapidly than the newly enriched gas can mix with the ambient medium. Steeper power-laws, $\alpha > 1$, develop when mixing occurs more rapidly than enrichment. If the ambient, primordial gas were truly infinite, the power-law would never completely encompass the delta function at $Z = Z_o$.

In reality, however, $M_o$ is not an inexhaustible reservoir of primordial gas. Eventually the entire ambient medium will become enriched to some $Z > Z_o$, and $p(Z)$ will consist entirely of a continuous, truncated power-law. As enrichment proceeds, gas near the low metallicity truncation of $p(Z)$, which still comprises much of the mass of the system, enriches towards higher metallicities in the power-law tail. This will produce a turnover at the low $Z$ limit of $p(Z)$. The low-turnover limit is produced by diffusion from many different sources, and is thus an additive process, which, as we noted above, tends to produce log-normal distributions. The physical interpretation of these two components is that the power-law tail represents newly enriched and poorly mixed gas that is above the average gas metallicity and is undergoing dilution, while the log-normal component represents the ambient medium that lies below the average gas metallicity and is undergoing enrichment.

>>>>

One-zone chemical evolution models assume homoegeneous mixing of metals from recent star formation into gas available for future star formation \citep[e.g.][]{Lanfranchi2006b,Kirby2011-III,Cote2017,Andrews2017}. While this assumes that metal mixing in the ISM plays no role in delaying future enrichment in star formation, accounting for this process is challening. More complex models \citep[e.g.][]{SchonrichBinney2009,Pezzulli2016} employ multiple (usually radial) zones or follow multiple gas phases to attempt to account for these effects. How to properly account for multi-phase mixing, however, is still poorly understood. This is in large part because both these models and large-scale cosmological simulations lack the necessary fidelity to capture the detailed, multi-phase mixing process of metals in the ISM directly. 
%Consequently, following metal mixing, if treated in these models, requires assumptions about how metals mix within the ISM and how metals couple to galactic outflows. 
Recent hydrodynamics simulations have employed parameteric models to account for unresolved sub-grid metal mixing \citep{PanScannapiecoScalo2013,Sarmento2017}, which plays an important role in determining the chemical properties of galaxies \citep[e.g.][]{Shen2010, Pilkington2012, Few2012, Brook2014, FengKrumholz2014, Escala2018, Armillotta2018} and the enrichment process and chemical signatures of the first stars \citep[e.g.][]{Jeon2015,Ritter2015,Smith2015}. Detailed hydrodynamics simulations incorporating a realistic multi-phase ISM and detailed stellar feedback are required to understand the details of metal mixing and metal outflows in galaxies.

In addition, it remains to be seen to what degree, if at all, metals of different enrichment origins (AGB winds, core collapse supernovae, Type Ia supernovae, neutron-star neutron-star mergers, etc.) may couple differently to the ISM. When metals are tracked individually, as opposed to a global metallicity field, their injection, mixing, and outflow properties are often treated uniformly. However, if metals do not behave uniformly in the ISM, if their mixing and ejection behavior depends directly upon the energetics and physical environment of their injection, then this assumption would need to be re-evaluated. Differences in metal distributions from ejecta in AGB stars, as compared to supernovae, has been explored recently in \cite{KrumholzTing2018}, but has yet to be demonstrated in hydrodynamics simulations.

The use of low mass dwarf galaxies, both observationally in the Local Group and in theoretical models, has been critical in improving our understanding of galactic chemical evolution. That dwarf galaxies efficiently pollute the circumgalactic medium and intergalactic medium (IGM) with metals has been demonstrated for some time both theoretically \citep[e.g.][]{DekelSilk1986,MacLowFerrara1999,Fragile2004,Muratov2017,Corlies2018}, and with direct observational evidence from the metal retention fractions of Local Group dwarf galaxies \citep[e.g.][]{Kirby2011-metals,McQuinn2015}. If there are differences in metal coupling to the ISM, this could potentially impact how metals are driven out of galaxies through galactic winds. This has implications for both interpreting observations of stellar abundances in nearby dwarf galaxies and in modeling galactic chemical evolution in both semi-analytic models and lower resolution cosmological simulations. However, examining this process has only become possible recently as it requires high resolution, galaxy-scale hydrodynamics simulations that can resolve ISM mixing and self-consistently drive galactic winds through multi-channel stellar feedback.
%It is commonly assumed that all metals couple equally to galactic outflows, however, which may not be a valid assumption if metals of different nucleosynthetic origins couple differently to the ISM \citep{KrumholzTing2018}. 
% Relaxing this assumption has implications for both interpreting observations of stellar abundances in nearby dwarf galaxies and in modeling galactic chemical evolution in both semi-analytic models and lower resolution cosmological simulations. However, testing this assumption has only become possible recently as it requires high resolution, galaxy-scale hydrodynamics simulations that can resolve ISM mixing and self-consistently drive galactic winds through multi-channel stellar feedback.

It is becoming increasingly valuable to develop a concrete theoretical understanding of galactic chemical evolution as a result of multiple, recent observational campaigns to probe detailed stellar abundances in our Milky Way and the Local Group, such as SEGUE \citep{Yanny2009}, RAVE \citep{Kunder2017}, APOGEE \citep{Anders2014}, and GALAH \citep{Buder2018}. Stellar abundances are directly imprinted with the enrichment pattern of their star forming cloud, whose chemical properties are determined by the process of turbulent metal mixing in the ISM. The degree to which we can associate "chemically tagged" stars as co-eval depends directly upon our understanding of metal enrichment and metal mixing in the ISM. Furthermore, this understanding is critical for using these observations to further deduce properties of a galaxy's evolutionary history. \textbf{One key aspect of this work has been a significant effort to characterize the number of independent dimensions accessible by the multi-dimensional chemical abundances observed in these studies \citep[e.g][]{Ting2012, Hogg2016,Jofre2017,Price-Jones2018}. Thus far, these studies have remained uninformed by hydrodynamics simulations and, conversely, these results cannot yet be used to constrain simulations. Doing so requires the kinds of high resolution, galaxy-scale multi-element chemodynamical simulations that have only become feasible recently.}

@@

\subsection{Physical Interpretation of the PDF}
\label{sec:interpretation}
Take the simple case of an initially primordial, uniform, isothermal medium of mass $M_o$ with initial metallicity $Z_{\rm o}$, containing a single, un-mixed enrichment event of mass $M_{\rm ej}$ whose size is small compared to the system and mass $M_{\rm ej} / M_o \ll 1$. Thus the background medium represents a virtually inexhaustible (but finite) source of un-enriched gas. In this case, $p(t,Z)$ initially takes the form of a double-delta function
\begin{equation}
p(t_{\rm o},Z) = \delta(Z_{\rm o}) + \frac{M_{\rm ej}}{M_o}\delta(Z - Z_{\rm ej}).
\end{equation}
If the enriched material mixes continually with the un-enriched gas at a constant rate, after some time this gas will have mixed with an equal amount of primordial gas \textbf{(lets take this time to be $\tau_{\rm mix}$). Therefore, we would have $p(\tau_{\rm mix},Z) = \delta(Z_{\rm o}) + \frac{2M_{\rm ej}}{M_o} \delta(Z - \frac{1}{2}Z_{\rm ej})$, $p(2\tau_{\rm mix},Z) = \delta(Z_{\rm o}) + \frac{4M_{\rm ej}}{M_o} \delta(Z - \frac{1}{4}Z_{\rm ej})$, and so on for other multiples of $\tau_{\rm mix}$.} This represents an inverse relationship between the mass of enriched gas and the metallicity of the enriched gas. \textbf{The superposition of the distributions of a single event evolving in time will appear as a power-law distribution with slope $\alpha = 1$.}

\textbf{Thus, if the gas is continually enriched by identical, well-separated enrichments events of mass $M_{\rm ej}$ and metallicity $Z_{\rm ej}$, the instantaneous $p(t,Z)$, or $p(Z)$, of these many events will be a power law in $Z$ with slope $\alpha = 1$, truncated at some minimim $Z$, related to the time since enrichment began, and some maximum, $Z_{\rm ej}$.}
% and thus a power-law evolution in time of the high metallicity portion of $p(Z)$ with slope $\alpha = 1$. 
The slope of the power-law, then, is determined by the rate of injection versus mixing with the ambient medium. A power-law index $\alpha < 1$ can occur when injection occurs more rapidly than the newly enriched gas can mix with the ambient medium. Steeper power-laws, $\alpha > 1$, develop when mixing occurs more rapidly than enrichment. If the ambient, primordial gas were truly infinite, the power-law would never completely encompass the delta function at $Z = Z_o$.

In reality, however, $M_o$ is not an inexhaustible reservoir of primordial gas. Eventually the entire ambient medium will become enriched to some $Z > Z_o$, and $p(Z)$ will consist entirely of a truncated power-law. As enrichment proceeds, gas near the low metallicity truncation of $p(Z)$, which still comprises much of the mass of the system, enriches towards higher metallicities in the power-law tail. This will produce a turnover at the low $Z$ limit of $p(Z)$. The low-turnover limit is produced by diffusion from many different sources, and is thus an additive process, which, as we noted above, tends to produce log-normal distributions. The physical interpretation of these two components is that the power-law tail represents newly enriched and poorly mixed gas that is above the average gas metallicity and is undergoing dilution, while the log-normal component represents the ambient medium that lies below the average gas metallicity and is undergoing enrichment.
